\chapter{Summary and outlook}

In this thesis, we aimed to analyze nuclear dynamics in higher excitations and considering propagation effects. Earlier work by Wolff and Evers \cite{wolff2023characterization} developed tools to characterize the low excitation regime (LER) and beyond LER. However, they did not account for the effects of light propagation through thick targets. Additionally, Burnham and Chiao solved the Maxwell-Bloch equations in an excitation regime spanning from LER to population inversion \cite{BURNHAM1969}, considering times much shorter than the characteristic time for spontaneous decay. In this thesis, we extended and generalized their approaches to higher excitations, including propagation effects.\\

First, we analyzed the nuclear dynamics and light propagation in the low-excitation regime for thin targets, where the effects of light propagation can be disregarded. Our results showed good correspondence with those of Wolff and Evers \cite{wolff2023characterization}. We demonstrated that the ratio of coherent to incoherent light intensity remains constant over time in the LER.\\

We then analyzed this ratio for thick targets, where propagation effects are significant. In this scenario, we observed notable deviations from the thin target limit, with the ratio no longer remaining constant over time.\\

Next, we extended our analysis to higher excitations, observing dynamics beyond LER, around population inversion, and beyond. In this regime, we discovered two interesting mechanisms: First, with increasing excitations, the minima of the coherent light intensity (Bessel beats) shift to later times. Second, a transition occurs around inversion, where the observed coherent light intensity shifts drastically and connects to the next minima.\\

This time shift effect of the observed coherent light depends on the target thickness. We proposed an experimental signature to detect non-linear excitations based on the relative time shift.\\

Finally, we analyzed nuclear dynamics using our model in the non-decaying limit ($\gamma = 0$), considering propagation effects at higher excitations. We found good agreement with the results of Burnham and Chiao \cite{BURNHAM1969}, concluding that spontaneous decay plays a crucial role in the creation of transitions observed around population inversion.

With the advancement of nuclear-resonant photons \cite{Shvydko2023,superradiance} and improved measurement schemes utilizing advanced focusing techniques, excitation beyond LER and around inversion is no longer infeasible. Our model can be used in future studies to model more complex systems, incorporating time- and position-dependent phases in the target, such as the dynamics in higher excitations compared to \cite{Shvydko2023}.

















































\iffalse 
In this thesis, we explored the time response of resonant nuclear scattering of x-rays beyond the low excitation regime. To achieve this, a general quantum optical formalism was derived, starting from the two-level systems, light-matter interaction, master equation, and Maxwell-Bloch equations. By applying the low excitation approximation, solving the Maxwell-Bloch equations, and Fourier transforming the response function from the frequency domain to the time domain, the time response was derived. An analogous calculation is provided in the appendix to also calculate $\rho_{eg}$ in the time domain within the low excitation regime (LER). Here we obtained that $\rho_{eg}$, which is proportional to the dipole moments in the target, is governed by a Bessel function of order 0, contrary to response function, which is governed by a Bessel function of order 1. \\

Going beyond the low excitation regime, the low excitation approximation can no longer be used, and the Maxwell-Bloch equations no longer decouple, making it impossible to obtain an analytical solution. To address this, we used numerical methods to study the propagation effects and nuclear dynamics. We implemented the method of lines (MOL) to discretize the spatial component of the Maxwell-Bloch equations, transforming the partial differential equation into $N$ coupled ordinary differential equations, which are much easier to solve.\\

After that, we demonstrated the capabilities of our numerical model by comparing the numerical and analytical solutions. Additionally, we showcased all the observables that can be analyzed using the model. We then introduced 2D excitation-time plots to better understand how the dynamics of our system evolve depending on the excitation. To begin, we scanned a low excitation regime, calculating for $\mathcal{A} = 0-0.012 \, \pi$. Using three block figures, we analyzed the ratios $|\rho_{eg}|^2/\rho_{ee}$ and $|\Omega|^2/\rho_{ee}$  in both the non-propagating limit and the thick target limit. We found good agreement with the results from \cite{wolff2023characterization} for both ratios in the non-propagating limit. We then analyzed these ratios in the propagation limit, generalizing the results from \cite{wolff2023characterization}.\\

We discovered that in the low-excitation regime, $\rho_{ee}$ and $\rho_{eg}$ exhibit the same dynamics, as they are coupled to each other. As the excitation level increases, the dynamics of $\rho_{eg}$ become suppressed in  $\rho_{ee}$, as the exponential decay starts to dominate in $\rho_{ee}$. These differing dynamics between $\rho_{ee}$ and $\rho_{eg}$ result in dips. To summarize, the ratio $|\rho_{eg}|^2/\rho_{ee}$, considering the propagation effects, remains constant only in ultra-low excitations and is generally not constant in thick targets. On the other, the ratio $|\Omega|^2/\rho_{ee}$ stays also not constant over time for thick targets, as $\Omega$ is governed by response function, which exhibits Bessel beats. $\rho_{ee}$ also shows bessel beats in the LER as a result of the coupling to $\rho_{eg}$. As a result, their ratio is not constant in the propagation limit. \\

We then calculated the summed incoherent radiation, as this is the observable which is experimentally accessible. Here analysed the ratio $I_\mathrm{coh}/ I_\mathrm{inc}$, where $I_\mathrm{inc}$ denotes the summed incoherent radiation $\sum_i\, \rho_{ee|i}$ and $I_\mathrm{coh}$ denotes the $|\Omega|^2$. The ratio is constant as a function of pulse area, but shows the characteristic Bessel beats from the $\Omega$. \\

As a next step, we scanned a bigger excitation regime, from $\mathcal{A} = 0\, \pi$ to $\mathcal{A} = 2\, \pi$. Observing $|\Omega|^2$, we discovered a very interesting \textit{avoided crossing like} structure at the population inversion, which we will be calling as the transition. In general, two interesting results are found beyond the low excitation regime. First, the Bessel beats shift in time with increasingly excitation. By population inversion, Bessel beats shift significantly and connect to the next Bessel beats after the inversion.\\


Other observables were also analyzed, such as $\rho_{ee}$ and $|\rho_{eg}|^2$. The dominance of the exponential decay in $\rho_{ee}$ became much more evident with increasing excitation levels. The summed incoherent light was calculated as well. According to the pulse area theorem, with a $\pi$ pulse, the population of excited states reaches its maximum, resulting in only incoherent radiation. With a $2\pi $ pulse, the population of excited states is zero, as are the generated dipole moments, which are proportional to $\rho_{eg}$. Therefore, no coherent or incoherent radiation occurs with a $2\pi $ pulse.\\

The strength of the time shift effect varies with increasing excitation, depending significantly on the target thickness. We fitted the position of the first Bessel beats in time for different target thicknesses and proposed an experimental signature to characterize nonlinear excitation based on the time shift of the Bessel beats.\\

Lastly, Burnham and Chiao \cite{BURNHAM1969} developed a model that approximates light propagation and nuclear dynamics up to the point of population inversion, considering times much shorter than the characteristic decay time. To replicate these conditions, we set $\gamma = 0$. We achieved good agreement with Burnham and Chiao's results in the non-decaying, propagating limit for excitations below population inversion.

\fi