\chapter*{Acknowledgements}

I want to thank all my colleagues at the Max Planck Institute for the wonderful times they have created over the last six months. Especially, I want to thank my office colleagues—Miriam Gerharz, Lukas Wolff, and Robert Horn—for the fruitful discussions, the casual chess games after lunch, and the occasional Turkish lessons. I also want to thank the group and Ze-an Peng for the joyful lunch talks.\\

Another thank you goes to Lukas for discussing his results with me and answering my questions, even when he was in a hurry to catch the bus in five minutes.\\

I also want to thank Miriam for co-supervising this thesis, patiently answering all my questions about programming and physics, and proofreading this thesis. And, of course, for helping me get started with Julia!\\

Last but not least, a very special thank you goes to Jörg Evers for welcoming me into his group and showing me how to do science from scratch. Without his guidance, this thesis would not have been possible. I am also very grateful for the opportunity to attend the DPG Meeting in Freiburg and present a poster for the first time in my life!\\

Bu tez aslında liseden sonra Almanya'da geçirdiğim ilk 4 yılın bitişini de simgeliyor ve bu noktada 'ekip'i anmazsam olmaz: Sarp, Sıla, Kaan, Beste ve Ceren. Acısıyla tatlısıyla son 4 yılda çok güzel vakitler geçirdik. Sickingenstrasse 23'ü de tepe tepe kullandık valla. Daha nice beraber yıllara.\\

Ve tabii ki sevgili annem ve babam. Maddi ve manevi her zaman arkamda olduğunuzu bilmek bana güç veriyor. Iyi ki varsınız!